### 2015-06-01 analyses/ResTime.Rmd

---
title: "How residence time constrains diversity"
author: "Ken Locey, Jay Lennon"
date: "May 31, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide'}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/QuantitativeBiodiversity/Assignments/Alpha") 
```

## INTRODUCTION

Here, we explore the constraining influence of ...  alpha ($\alpha$) diversity. 
We will quantify general components of $\alpha$-diversity: **richness** and **evenness**.
From there, we ...  **species-abundance distributions (SAD)**.

```{r, results = 'hide', message=FALSE}
#install.packages("vegan") 
#instead, use try("vegan")
require("vegan") 
```

## 2) LOADING DATA

```{r}
sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/SimData.csv")  
```

```{r}
par(mfrow=c(2, 2))
x <- log(sim.data$TracerParticle_ResTime)
N <- log(sim.data$TotalAbundance)
plot(x, N, xlab="Particle Residence Time", 
     ylab="Total Abundance", 
     main = "N vs. Tau", col='SteelBlue')

x <- log(sim.data$MicrobeCell_ResTime)
plot(x, N, xlab="Cell Residence Time", 
     ylab="Total Abundance", 
     main = "N vs. Tau", col='darkorchid4')

x <- log(sim.data$TracerParticle_ResTime)
S <- log(sim.data$SpeciesRichness)
plot(x, S, xlab="Particle Residence Time", 
     ylab="Total Abundance", 
     main = "S vs. Tau", col='SteelBlue')

x <- log(sim.data$MicrobeCell_ResTime)
plot(x, S, xlab="Cell Residence Time", 
     ylab="Total Abundance", 
     main = "S vs. Tau", col='darkorchid4')
```

```{r}
rads <- c()
rad.list <- readLines("~/GitHub/hydrobide/results/simulated_data/RADs.csv")  
for (rad in rad.list) {
  rad <- unlist(strsplit(rad, split = ","))  
  rad <- as.numeric(rad)
  rads <- c(rads, rad)
  }
```