### 2015-06-01 analyses/ResTime.Rmd

---
title: "How residence time constrains diversity"
author: "Ken Locey, Jay Lennon"
date: "May 31, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/hydrobide") 
```

## INTRODUCTION

Here, we explore the constraining influence of ...  alpha ($\alpha$) diversity. 
We will quantify general components of $\alpha$-diversity: **richness** and **evenness**.
From there, we ...  **species-abundance distributions (SAD)**.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan")

sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/SimData.csv")  
```

## Univariate relationships

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
par(mfrow=c(1, 2), width = 8, height = 8)

height <- sim.data$Height
width <- sim.data$Width
flow.rate <- sim.data$FlowRate

h.tau <- log((height * width)/flow.rate)
m.tau <- log(sim.data$MicrobeCell_ResTime)
t.tau <- log(sim.data$TracerParticle_ResTime)

plot(h.tau, m.tau, xlab="Hydraulic Residence Time", 
     ylab="Cell Residence Time", 
     col='SteelBlue')

plot(h.tau, t.tau, xlab="Hydraulic Residence Time", 
     ylab="Particle Residence Time", 
     col='darkorchid4')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
par(mfrow=c(2, 2))

N <- log(sim.data$TotalAbundance)
S <- log(sim.data$SpeciesRichness)

plot(t.tau, N, xlab="Particle Residence Time", 
     ylab="Total Abundance", 
     main = "N vs. Tau", col='SteelBlue')

plot(m.tau, N, xlab="Cell Residence Time", 
     ylab="Total Abundance", 
     main = "N vs. Tau", col='darkorchid4')

plot(t.tau, S, xlab="Particle Residence Time", 
     ylab="Species Richness", 
     main = "S vs. Tau", col='SteelBlue')

plot(m.tau, S, xlab="Cell Residence Time", 
     ylab="Species Richness", 
     main = "S vs. Tau", col='darkorchid4')

```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
par(mfrow=c(2, 2))

N <- sim.data$TotalAbundance
plot(m.tau, log(N), xlab="Cell Residence Time", 
     ylab="Total Abundance", 
     main = "N vs. Tau",
     col='SteelBlue')

S <- sim.data$SpeciesRichness
plot(m.tau, log(S), xlab="Cell Residence Time", 
     ylab="Species Richness", 
     main = "S vs. Tau",
     col='darkorchid4')

E <- sim.data$Evar
plot(m.tau, E, xlab="Cell Residence Time", 
     ylab="Species Evenness", 
     main = "Evenness vs. Tau",
     col='SteelBlue')

avg.ab <- N/S
plot(m.tau, log(avg.ab), xlab="Cell Residence Time", 
     ylab="Average Abundance",
     main = "N/S vs. Tau",
     col='darkorchid4')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
par(mfrow=c(2, 2))

BP <- sim.data$BergerParker
plot(m.tau, BP, xlab="Cell Residence Time", 
     ylab="Berger-Parker", 
     main = "Relative Dominance",
     col='SteelBlue')

SimpD <- sim.data$InvSimpDiversity
plot(m.tau, SimpD, xlab="Cell Residence Time", 
     ylab="Simpsons, 1/D", 
     main = "Diversity",
     col='darkorchid4')

Skew <- sim.data$Skew
plot(m.tau, Skew, xlab="Cell Residence Time", 
     ylab="Skewness", 
     main = "Rarity",
     col='SteelBlue')

Nmax <- sim.data$Nmax
plot(m.tau, log(Nmax), xlab="Cell Residence Time", 
     ylab="Nmax",
     main = "Absolute Dominance",
     col='darkorchid4')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
par(mfrow=c(2, 2))

plot(log(N), log(S), xlab="Total abundance",
     ylab="Species Richness", 
     main = "S vs. N",
     col='SteelBlue')

plot(log(N), log(Nmax), xlab="Total abundance", 
     ylab="Nmax", 
     main = "Nmax vs. N",
     col='darkorchid4')

plot(log(N), log(E), xlab="Total abundance", 
     ylab="Species Evenness", 
     main = "Evenness vs. N",
     col='SteelBlue')

plot(log(N), log(Skew), xlab="Total abundance", 
     ylab="Skewness",
     main = "Rarity vs. N",
     col='darkorchid4')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rads <- c()
rad.list <- readLines("~/GitHub/hydrobide/results/simulated_data/RADs.csv")  
for (rad in rad.list) {
  rad <- unlist(strsplit(rad, split = ","))  
  rad <- as.numeric(rad)
  rads <- c(rads, rad)
  }
```