### 2015-06-17 analyses/2015_06_16_ResTime_GeneralResults.Rmd
modified Rmd file, generated new plots and .pdf, modified bide and model functions -KL

---
title: "How residence time constrains diversity"
author: "Ken Locey, Jay Lennon"
date: "May 31, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/hydrobide") 
```

## INTRODUCTION

Here, we explore the influence of mean cell residence time, hydraulic residence time, and particle residence time on abundance, local ($\alpha$) diversity, physiology, and resource availability and diversity.

These simulations

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
```

```{r, results = 'hide', message = FALSE, warning = FALSE}
## Load Data
sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015June/June16/SimData.csv")  
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Assign Data
sim <- sim.data$sim # the specific simulation run
height <- sim.data$Height
width <- sim.data$Width
flow.rate <- sim.data$FlowRate
seed <- sim.data$starting.seed

h.tau <- log((height * width)/flow.rate)
m.tau <- log(sim.data$individual.tau)
t.tau <- log(sim.data$particle.tau)

N <- sim.data$N
S <- sim.data$S
e.var <- sim.data$e.var
e.simp <- sim.data$simpson.e
avg.ab <- N/S
Skew <-sim.data$skew
Nmax <- sim.data$N.max
SimpD <- sim.data$SD

res.inflow <- sim.data$res.inflow
res.types <- sim.data$res.types
max.res.val <- sim.data$max.res.val

res.conc <- sim.data$resource.concentration
num.tracers <- sim.data$tracer.particles
num.resources <- sim.data$resource.particles
```

## Univariate relationships

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
par(mfrow=c(1, 2))

num.sims <- max(sim.data$sim)
colors <- rainbow(num.sims)

i <- 0
while (i < num.sims){

  sim.sub <- subset(sim.data, sim == i)
  i <- i + 1
  print(i)

  height <- sim.sub$Height
  width <- sim.sub$Width
  flow.rate <- sim.sub$FlowRate
  e.var <- sim.sub$e.var
  e.simp <- sim.sub$simpson.e
  avg.ab <- N/S
  Skew <-sim.sub$skew
  Nmax <- sim.sub$N.max
  SimpD <- sim.sub$SD

  h.tau <- log((height * width)/flow.rate)

  N <- sim.sub$N
  N <- log(N)
  S <- sim.sub$S
  S <- log(S)

  x <- c()
  y <- c()
  ct <- 1
  for (val in N){
    if(is.nan(val) == FALSE && val > 0.1){
      y <- c(y, val)
      x <- c(x, h.tau[ct])
      }
    ct <- ct+1
    }

  if (i == 1){
    plot(x, y, xlab="log(ecosystem residence time)", 
    ylab="Total Abundance", ylim=c(3.5, 10), xlim=c(3, 9),
    main = "N vs. Tau", col=colors[i])
    lines(x, y, col=colors[i])
    }
  else{
    points(x, y, col=colors[i])
    lines(x, y, col=colors[i])
    }

  }

i <- 0
while (i < num.sims){

  sim.sub <- subset(sim.data, sim == i)
  i <- i + 1

  height <- sim.sub$Height
  width <- sim.sub$Width
  flow.rate <- sim.sub$FlowRate
  e.var <- sim.sub$e.var
  e.simp <- sim.sub$simpson.e
  avg.ab <- N/S
  Skew <-sim.sub$skew
  Nmax <- sim.sub$N.max
  SimpD <- sim.sub$SD

  h.tau <- log((height * width)/flow.rate)
  S <- sim.sub$S
  S <- log(S)

  x <- c()
  y <- c()
  ct <- 1
  for (val in S){
    if(is.nan(val) == FALSE && val > 0.1){
      y <- c(y, val)
      x <- c(x, h.tau[ct])
      }
    ct <- ct+1
    }

  if (i == 1){
    plot(x, y, xlab="log(ecosystem residence time)", 
    ylab="log(species richness)", ylim=c(0, 4), xlim=c(3, 9),
    main = "S vs. Tau", col=colors[i])
    lines(x, y, col=colors[i])
    }
  else{
    points(x, y, col=colors[i])
    lines(x, y, col=colors[i])
    }
}
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Reassign Data

sim <- sim.data$sim # the specific simulation run
height <- sim.data$Height
width <- sim.data$Width
flow.rate <- sim.data$FlowRate
seed <- sim.data$starting.seed

h.tau <- log((height * width)/flow.rate)
m.tau <- log(sim.data$individual.tau)
t.tau <- log(sim.data$particle.tau)

N <- sim.data$N
S <- sim.data$S
e.var <- sim.data$e.var
e.simp <- sim.data$simpson.e
avg.ab <- N/S
Skew <-sim.data$skew
Nmax <- sim.data$N.max
SimpD <- sim.data$SD

res.inflow <- sim.data$res.inflow
res.types <- sim.data$res.types
max.res.val <- sim.data$max.res.val

res.conc <- sim.data$resource.concentration
num.tracers <- sim.data$tracer.particles
num.resources <- sim.data$resource.particles
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
par(mfrow=c(1, 2))

growth <- sim.data$avg.per.capita.growth
x <- c()
y <- c()

ct <- 1
for (val in growth){
  if(is.nan(val) == FALSE && val > 0.1){
    y <- c(y, val)
    x <- c(x, h.tau[ct])
    }
  ct <- ct+1
  }

plot(x, y, xlab="log(ecosystem residence time)", 
     ylab="growth rate", 
     main = "Per capita growth vs. Tau",
     col='dodgerblue')

maint <- (S*sim.data$avg.per.capita.maint)/N
plot(h.tau, log(maint), xlab="log(ecosystem residence time)", 
     ylab="log(maintenance)", 
     main = "Per capita maintenance vs. Tau",
     col='darkolivegreen')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
par(mfrow=c(2, 2))

plot(log(N), Skew, xlab="Total abundance", 
     ylab="Skewness",
     main = "Rarity vs. N",
     col='plum')

plot(log(N), log(Nmax), xlab="Total abundance", 
     ylab="Nmax", 
     main = "Nmax vs. N",
     col='darkorange')

plot(log(N), log(e.var), xlab="Total abundance", 
     ylab="Species Evenness", 
     main = "Evenness vs. N",
     col='lightseagreen')

plot(log(N), log(S), xlab="Total abundance",
     ylab="Species Richness", 
     main = "S vs. N",
     col='indianred')

```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
par(mfrow=c(1, 2))

xlabel <- "log(ecosystem residence time)"
res.conc <- sim.data$resource.concentration
res.div <- sim.data$shannons.resource.diversity
res.rich <- sim.data$resource.richness

x <- c()
y <- c()
ct <- 1
for (val in res.conc){
  if(is.nan(val) == FALSE && val > 0.1){
    y <- c(y, val)
    x <- c(x, h.tau[ct])
    }
  ct <- ct+1
  }

plot(x, log(y), xlab=xlabel, 
     ylab="density", 
     main = "Resource density vs. Tau",
     col='peru')

x <- c()
y <- c()
ct <- 1
for (val in res.rich){
  if(is.nan(val) == FALSE && val > 0.1){
    y <- c(y, val)
    x <- c(x, h.tau[ct])
    }
  ct <- ct+1
  }

plot(x, y, xlab=xlabel, 
     ylab="Number of resource types", 
     main = "Resource richness vs. Tau",
     col='slateblue')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rads <- list()
rad.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015June/June16/RADs.csv",
            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
            as.is = TRUE, skipNul = TRUE, fill = TRUE)

rad.list[is.na(rad.list)] <- 0
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
species <- list()
species.list <- read.table("~/GitHub/hydrobide/results/simulated_data/2015June/June16/Species.csv",
            header = FALSE, sep = ",", col.names = paste0("V", seq_len(1000)),
            as.is = TRUE, skipNul = TRUE, fill = TRUE)

species.list[is.na(species.list)] <- -1
```