### 2015-08-26 analyses/2015_06_27_ResTime.Rmd

---
title: "Residence time, a master variable of ecology, physiology, and natural selection"
author: "Ken Locey, Jay Lennon"
date: "June, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/hydrobide") 
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
#install.packages("moments")
require("moments")
library(ggplot2)
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Load Data
sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/12_Aug/SimData.csv")  
#sim.data <- subset(sim.data, motion == ' random_walk ')
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Assign Data
sim <- sim.data$sim # the specific simulation run
height <- sim.data$Height
width <- sim.data$Width
flow.rate <- sim.data$FlowRate
seed <- sim.data$starting.seed

h.tau <- (height * width)/flow.rate
m.tau <- sim.data$individual.tau
t.tau <- sim.data$particle.tau
r.tau <- sim.data$resource.tau

N <- sim.data$N
S <- sim.data$S
e.var <- sim.data$e.var
e.simp <- sim.data$simpson.e
avg.ab <- N/S
Skew <-sim.data$skew
Nmax <- sim.data$N.max
SimpD <- sim.data$SD
Whittaker.turnover <- sim.data$species.turnover

res.inflow <- sim.data$res.inflow
res.types <- sim.data$res.types
max.res.val <- sim.data$max.res.val

res.conc <- sim.data$resource.concentration
num.tracers <- sim.data$tracer.particles
num.resources <- sim.data$resource.particles
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Loading files with plotting and diversity functions
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
source("~/GitHub/hydrobide/tools/Rbin/metrics.R")
```

## We expected mean residence times for individuals, inert tracers, and resource particles to reflect ecosystem residence time.

## We hypothesized that total abundance (N) and species richness (S) shared humped-shaped relationshiped with residence time, reflecting chemostat theory.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}

source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")

png(filename = '~/GitHub/hydrobide/results/figs/NandS_vs_ResTime.png',
      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")

num.sims <- max(sim.data$sim)
colors <- rainbow(num.sims)

sim.data$N[sim.data$N<=0] <- 1
sim.data$N[sim.data$S<=0] <- 1

p1 <- ggplot(sim.data, aes(x=h.tau, y=N, group=sim, colour=sim)) +
  scale_color_gradient(high="cyan") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  theme_black() +
  labs(x= expression(tau), y="Individuals") +
  labs(title="Total abundance") +
  guides(colour=FALSE) + 
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = .48)

p2 <- ggplot(sim.data, aes(x=h.tau, y=S, group=sim, colour=sim)) +
  scale_color_gradient(high="chartreuse") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  theme_black() +
  labs(x= expression(tau), y="Number of species") +
  labs(title="Species richness") +
  guides(colour=FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 0.8)

multiplot(p1, p2, cols=2)
dev.off()
```

## We hypothesized that the realtionship species evenness (E) to residence time would be U-shaped, and that the relationship of species turnover (B) to residence time would be L-shaped.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}

png(filename = '~/GitHub/hydrobide/results/figs/EandB_vs_ResTime.png',
      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")

p1 <- ggplot(sim.data, aes(x=h.tau, y=e.var, group=sim, colour=sim)) +
  scale_color_gradient(high="cyan") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  theme_black() +
  labs(x= expression(tau), y="Evar") +
  labs(title="Evenness") +
  guides(colour=FALSE) + 
  scale_x_log10() + 
  coord_fixed(ratio = 1.8)

p2 <- ggplot(sim.data, aes(x=h.tau, y=Whittaker.turnover, group=sim, colour=sim)) +
  scale_color_gradient(high="chartreuse") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  theme_black() +
  labs(x= expression(tau), y="Whittaker's") +
  labs(title="Species turnover") +
  guides(colour=FALSE) +
  scale_x_log10() +
  coord_fixed(ratio = 2.0)

multiplot(p1, p2, cols=2)
dev.off()
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Reassign Data

sim <- sim.data$sim # the specific simulation run
height <- sim.data$Height
width <- sim.data$Width
flow.rate <- sim.data$FlowRate
seed <- sim.data$starting.seed

h.tau <- (height * width)/flow.rate
m.tau <- sim.data$individual.tau
t.tau <- sim.data$particle.tau

N <- sim.data$N
S <- sim.data$S
e.var <- sim.data$e.var
e.simp <- sim.data$simpson.e
avg.ab <- N/S
Skew <-sim.data$skew
Nmax <- sim.data$N.max
SimpD <- sim.data$SD

res.inflow <- sim.data$res.inflow
res.types <- sim.data$res.types
max.res.val <- sim.data$max.res.val

res.conc <- sim.data$resource.concentration
num.tracers <- sim.data$tracer.particles
num.resources <- sim.data$resource.particles
```

## We expected lower maintenance requirements at higher residence times, where individuals could grow quickly, scrambling to consume available resources, or have an overall slow life history.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}

png(filename = '~/GitHub/hydrobide/results/figs/Growth_and_Maint.png',
      width = 700, height = 500, units = "px", pointsize = 17, bg = "black")

growth <- sim.data$avg.per.capita.growth
xs <- c()
ys <- c()

ct <- 1
for (val in growth){
  if(is.nan(val) == FALSE && val > 0.1){
    ys <- c(ys, val)
    xs <- c(xs, h.tau[ct])
    }
  ct <- ct+1
  }

p1 <- ggplot(data=NULL, aes(x=xs, y=ys)) +
  geom_point(colour="magenta") +
  labs(x= expression(tau), y="Growth") +
  labs(title="Growth rate") +
  scale_x_log10() +
  coord_fixed(ratio = 5.3) +
  theme_black()

maint <- (S*sim.data$avg.per.capita.maint)/N

p2 <- ggplot(data=NULL, aes(x=h.tau, y=maint)) +
  geom_point(colour="cyan") +
  labs(x= expression(tau), y="Maintenance") +
  labs(title="Maintenance") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 0.34) +
  theme_black()

multiplot(p1, p2, cols=2)
dev.off()
```

## We hypothesized that aspects of diversity would be related to total abundance (N), reflecting the feasible set and the influence of residence time.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
png(filename = '~/GitHub/hydrobide/results/figs/Diversity_vs_N.png',
      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")

par(mfrow=c(2, 2))

plot(N, Skew, xlab="Total abundance", log="xy",
     ylab="Skewness",
     main = "Rarity vs. N",
     col='plum')

plot(N, Nmax, xlab="Total abundance", log="xy",
     ylab="Nmax", 
     main = "Nmax vs. N",
     col='darkorange')

plot(N, e.var, xlab="Total abundance", log="xy",
     ylab="Species Evenness", 
     main = "Evenness vs. N",
     col='lightseagreen')

plot(N, S, xlab="Total abundance", log="xy",
     ylab="Species Richness", 
     main = "S vs. N",
     col='indianred')

dev.off()
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Loading simulated data csv files

sims <- c()
cts <- c()
tracer.RTDs <- list()
res.RTDs <- list()
ind.RTDs <- list()

inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/IndRTD.csv"
con  <- file(inputFile, open = "r")
ct = 1
while (length(myLine <- scan(con, what="numeric", nlines=1, sep=',', skip=0, quiet=TRUE)) > 0 ){
  line <- as.numeric(myLine)
  cts <- c(cts, line[1])
  sims <- c(sims, line[2])
  rtd <- line[-1:-2]
  ind.RTDs[[ct]] <- rtd
  ct <- ct + 1
  } 
close(con)

inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/TracerRTD.csv"
con  <- file(inputFile, open = "r")
ct = 1
while (length(myLine <- scan(con, what="numeric", nlines=1, sep=',', skip=0, quiet=TRUE)) > 0 ){
  line <- as.numeric(myLine)
  rtd <- line[-1:-2]
  tracer.RTDs[[ct]] <- rtd
  ct <- ct + 1
  } 
close(con)

inputFile <- "~/GitHub/hydrobide/results/simulated_data/2015_June/26_June/ResRTD.csv"
con  <- file(inputFile, open = "r")
ct = 1
while (length(myLine <- scan(con, what="numeric", nlines=1, sep=',', skip=0, quiet=TRUE)) > 0 ){
  line <- as.numeric(myLine)
  rtd <- line[-1:-2]
  res.RTDs[[ct]] <- rtd
  ct <- ct + 1
  } 
close(con)
```

# We hypothesized that, as often observed for tracers in real systems, residence time distributions (RTDs) for individuals and resource particles would be lognormal-like curves, which appear normal or left-skewed on a log x-axis. Colors correspond across plots.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
png(filename = '~/GitHub/hydrobide/results/figs/RTDs.png',
      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")

par(mfrow=c(2, 2))

t.skews <- c()
i.skews <- c()
r.skews <- c()
t.means <- c()
i.means <- c()
r.means <- c()

colors <- rainbow(length(ind.RTDs))
plot(density(tracer.RTDs[[1]]), col = 'white', xlab='Residence Time',
     xlim = c(0,8), ylim = c(0,2),
     ylab='Probability Density',  main = "Tracer RTD")

ct <- 1
for(rtd in tracer.RTDs){
  if(any(is.na(rtd)) == FALSE){
    lines(density(log(rtd), bw = 0.2), col = colors[ct])
    t.skews <- c(t.skews, skewness(rtd))
    t.means <- c(t.means, mean(rtd))
    }
  ct = ct + 1
}

plot(density(ind.RTDs[[1]]), col = 'white', xlab='Residence time',
  xlim = c(0,8), ylim = c(0,2),
  ylab='Probability Density', main = "Individual RTD")

ct <- 1
for(rtd in ind.RTDs){
  if(any(is.na(rtd)) == FALSE){
    lines(density(log(rtd), bw = 0.2), col = colors[ct])
    i.skews <- c(i.skews, skewness(rtd))
    i.means <- c(i.means, mean(rtd))
    }
  ct = ct + 1
}

#plot(density(res.RTDs[[1]]), col = 'white', xlab='Residence time',
#     xlim = c(0,8), ylim = c(0,2),
#     ylab='Probability Density', main = "Resource RTD")

#ct <- 1
#for(rtd in res.RTDs){
#  if(any(is.na(rtd)) == FALSE){
#    lines(density(log(rtd), bw = 0.2), col = colors[ct])
#    r.skews <- c(r.skews, skewness(rtd))
#    r.means <- c(r.means, mean(rtd))
#    }
#  ct = ct + 1
#}

dev.off()
```

# Increasing ecosystem residence time shifts the mean of the RTD to higher average values (older individuals). But, it also changes the skewness of the RTD from asymmetrically young to asymmetrically old. No significant effect is seen among tracers.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=6, fig.height=6}

png(filename = '~/GitHub/hydrobide/results/figs/RTD_shifts.png',
      width = 700, height = 500, units = "px", pointsize = 17, bg = "white")

par(mfrow=c(1, 1))

plot(log(t.means), t.skews, col = 'grey', xlab='Mean of the RTD, i.e., avg. residence time', ylab='Skewness of the RTD', lwd=2, xlim = c(0,7), ylim = c(-1,2),
     main = "Increasing mean residence time flips the age
     structure of the community from young to old")

points(log(i.means), i.skews, lwd=2, col = 'darkmagenta')
#points(r.means, r.skews, lwd=2, col = 'darkcyan')
lines(c(0,9), c(0,0), lty=2, col='black')

i.OLS <- lm(i.skews ~ log(i.means))
coeff <- summary(i.OLS, correlation = TRUE)
i.slope <- round(coefficients(i.OLS)[2], 3)
i.p <- round(summary(i.OLS)$coefficients[8], 3)
abline(i.OLS, col = "darkmagenta", lw = 2)

t.OLS <- lm(t.skews ~ log(t.means))
coeff <- summary(t.OLS, correlation = TRUE)
t.slope <- round(coefficients(t.OLS)[2], 2)
t.p <- round(summary(t.OLS)$coefficients[8], 2)
abline(t.OLS, col = "grey", lw = 2)

legend("topright", legend = c(paste("Tracers: slope=",t.slope,", p=",t.p), paste("Individuals: slope=",i.slope,", p=",i.p)), pch=19, bty = "n", col = c("grey", "darkmagenta"))

dev.off()
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=6, fig.height=6}
num.sims <- max(sim.data$sim) + 1

i <- 0
while (i < num.sims){

  sp.sub <- subset(sim.data$species.list, sim == i)
  sp.beta <- c()
  j <- 2
  while(j <= length(sp.sub)){
    print(j)
    w.beta <- sp.turnover(sp.sub[j], sp.sub[j-1])
    sp.beta <- c(sp.beta, w.beta)
    }
    j <- j+1
    }
  i <- i+1
  print(sp.beta)
  }
```