### 2015-09-09 analyses/ForPaper/MultivariateTests.Rmd

---
title: "Residence time, a master variable of ecology, physiology, and natural selection"
author: "Ken Locey, Jay Lennon"
date: "June, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/hydrobide") 
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
require("picante")
require("ape")
require("fossil")
require("simba")
require("reshape")

#install.packages("vegan") 
require("vegan") 
#install.packages("moments")
require("moments")
library(ggplot2)
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Loading files with plotting and diversity functions
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
source("~/GitHub/hydrobide/tools/Rbin/metrics.R")
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Load Data
sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/29_Aug/SimData.csv")  
sim.data$h.tau <- (sim.data$height * sim.data$width)/sim.data$flow.rate
sim.data$total.abundance[sim.data$total.abundance<=0] <- 1
sim.data$species.richness[sim.data$species.richness<=0] <- 1

sim.data[is.na(sim.data)] <- 0
```

## We hypothesized that total abundance (N) and species richness (S) shared humped-shaped relationshiped with residence time, reflecting chemostat theory.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/NandS_vs_ResTime.png',
      width = 1500, height = 1200, res=200, bg = "white")
```

## 7.) Variance partitioning: 
```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=4}

# Physical and metacommunity variables
phys.dat <- as.matrix(subset(sim.data, 
                      select = c(h.tau)))

# Resource variables
res.dat <- as.matrix(subset(sim.data, 
                      select = c(resource.concentration,
                      shannons.resource.diversity,
                      resource.richness,
                      resource.particles,
                      avg.per.capita.growth,
                      avg.per.capita.maint,
                      avg.per.capita.N.efficiency,
                      avg.per.capita.P.efficiency,
                      avg.per.capita.C.efficiency,
                      avg.per.capita.active.dispersal)))

y.dat <- as.matrix(subset(sim.data, 
                      select = c(total.abundance,
                      species.richness,
                      simpson.e,
                      species.turnover)))

rda.phys <- rda(y.dat ~ phys.dat)
#rda.phys
rda.res <- rda(y.dat ~ res.dat)
#rda.res

# Two explanatory matrices -- Hellinger-transform Y
mod <- varpart(Y=y.dat, X=phys.dat, res.dat,  transfo="log")
mod
showvarparts(2)
plot(mod)
mtext("Variation partitioning")
text(1.6, 0.6, "env") 
text(-0.6, 0.6, "tau") 
```

```{r, results='hide', echo=F, message=F, warning=F, fig.width=7, fig.height=4}

# Test fraction using RDA:
phys.anova <- anova(rda.phys, step=200, perm.max=200)
phys.anova

res.anova <- anova(rda.res, step=200, perm.max=200)
res.anova

bio.anova <- anova(rda.bio, step=200, perm.max=200)
bio.anova
```

### ii.  Constrained Ordination
Another way we can test hypotheses with continuous data is to use **constrained ordination**, which is sometimes referred to as canonical ordination.
Constrained ordination explores the relationships between two matrices: an **explanatory matrix** and a **response matrix**. 
Canonical correspondence analysis (CCA) and redundancy analysis (RDA) are two types of constrained ordination.
These techniques are based on the linear model framework and thus can be used to formally test hypotheses.
Constrained ordination works by first conducting multivariate multiple linear regression followed either by correspondence analysis (CA) with CCA or Principal Components Analysis (PCA) with RDA, while using the matrix of fitted values to obtain a constrained ordination.
A permutation test can then be used to test for overall significance.

Here, we will use environmental data to conduct a CCA on the fish assemblages of the Doubs River. 
We will start by creating an explanatory matrix that contains water chemistry data.
We will then use the `cca()` function from the `vegan` package.
Note, we have to specify that we want the `cca` function in the `vegan` package because there are `cca` functions in both `vegan` and `ade4`!
We will then use permutation tests to evaluate the significance of our model. 
Finally, we will test the influence of each environmental variable on the constrained axes.

```{r, results = "hide"}
# Get Dominant members

# Define Environmental Matrix
#bio.dat
#phys.dat
#res.dat
all.dat <- cbind(bio.dat, phys.dat, res.dat)
# Conduct CCA 
y.cca <- vegan::cca(y.dat ~ all.dat)

# Permutation Tests
anova(y.cca, by = "axis")
cca.fit <- envfit(y.cca, all.dat, perm = 999)
cca.fit

# Calculate Explained Variation
cca.explainvar1 <- round(y.cca$CCA$eig[1] / 
                         sum(c(y.cca$CCA$eig, y.cca$CA$eig)), 3) * 100

cca.explainvar2 <- round(y.cca$CCA$eig[2] / 
                         sum(c(y.cca$CCA$eig, y.cca$CA$eig)), 3) * 100

# Define Plot Parameters
par(mar = c(5, 5, 4, 4) + 0.1)

# Initiate Plot
plot(scores(y.cca, display = "wa"), xlim = c(-3.5, 2), ylim = c(-3.2, 3.2),
     xlab = paste("CCA 1 (", cca.explainvar1, "%)", sep = ""),
     ylab = paste("CCA 2 (", cca.explainvar2, "%)", sep = ""),
     pch = 16, cex = 2.0, type = "n", cex.lab = 1.5,
     cex.axis = 1.2, axes = FALSE)

# Add Axes
axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 1.2, las = 1)
abline(h = 0, v = 0, lty = 3)
box(lwd = 2)

# Add Points & Labels
points(scores(y.cca, display = "wa"),
       pch = 19, cex = 3, bg = "gray", col = "gray")
text(scores(y.cca, display = "wa"), 
     labels = row.names(scores(y.cca, display = "wa")))

# Add Environmental Vectors
vectors <- scores(active.cca, display = "bp")
row.names(vectors) <- c("Depth", "Temp", "pH","DOC", "DON", "TDS", "Salinity")
arrows(0, 0, vectors[,1] * 2, vectors[, 2] * 2, 
       lwd = 2, lty = 1, length = 0.2, col = "red")
text(vectors[,1] * 2, vectors[, 2] * 2, pos = 3, 
     labels = row.names(vectors))
axis(side = 3, lwd.ticks=2, cex.axis=1.2, las = 1, col = "red", lwd = 2.2,
     at = pretty(range(vectors[, 1])) * 2, labels = pretty(range(vectors[, 1])))
axis(side = 4, lwd.ticks=2, cex.axis=1.2, las = 1, col = "red", lwd = 2.2,
     at = pretty(range(vectors[, 2])) * 2, labels = pretty(range(vectors[, 2])))
```