### 2015-09-09 analyses/ForPaper/PrimaryPredictions.Rmd

---
title: "Residence time, a master variable of ecology, physiology, and natural selection"
author: "Ken Locey, Jay Lennon"
date: "June, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/hydrobide") 
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
#install.packages("moments")
require("moments")
library(ggplot2)
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Loading files with plotting and diversity functions
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
source("~/GitHub/hydrobide/tools/Rbin/metrics.R")
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Load Data
sim.data <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_August/29_Aug/SimData.csv")  
h.tau <- (sim.data$height * sim.data$width)/sim.data$flow.rate
sim.data$total.abundance[sim.data$total.abundance<=0] <- 1
sim.data$species.richness[sim.data$species.richness<=0] <- 1

sim.data.fluid <- subset(sim.data, motion == "fluid")
h.tau.fluid <- (sim.data.fluid$height * sim.data.fluid$width)/sim.data.fluid$flow.rate
sim.data.fluid$total.abundance[sim.data.fluid$total.abundance<=0] <- 1
sim.data.fluid$species.richness[sim.data.fluid$species.richness<=0] <- 1

sim.data.random_walk <- subset(sim.data, motion == "random_walk")
h.tau.random_walk <- (sim.data.random_walk$height * sim.data.random_walk$width)/sim.data.random_walk$flow.rate
sim.data.random_walk$total.abundance[sim.data.random_walk$total.abundance<=0] <- 1
sim.data.random_walk$species.richness[sim.data.random_walk$species.richness<=0] <- 1
```

## We hypothesized that total abundance (N) and species richness (S) shared humped-shaped relationshiped with residence time, reflecting chemostat theory.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/NandS_vs_ResTime.png',
      width = 1500, height = 1200, res=200, bg = "white")

p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, 
                           group=sim.data$sim, colour=sim.data$sim)) +
  scale_color_gradient(high="cyan") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="Individuals") +
  labs(title="Total abundance") +
  guides(colour=FALSE) + 
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = .50)

p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, 
                           group=sim.data$sim, colour=sim.data$sim)) +
  scale_color_gradient(high="chartreuse") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="Number of species") +
  labs(title="Species richness") +
  guides(colour=FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.0)

multiplot(p1, p2, cols=2)
dev.off()
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/NandS_vs_ResTime_ByMotion.png',
    width = 1500, height = 1200, res=200, bg = "white")

p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$total.abundance, 
                           group=sim.data$sim, colour=sim.data$motion)) +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="Individuals") +
  labs(title="Total abundance") +
  guides(colour=FALSE) + 
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = .50)

p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.richness, 
                           group=sim.data$sim, colour=sim.data$motion)) +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="Number of species") +
  labs(title="Species richness") +
  guides(colour=FALSE) +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.0)

multiplot(p1, p2, cols=2)
dev.off()
```

## We hypothesized that the realtionship species evenness (E) to residence time would be U-shaped, and that the relationship of species turnover (B) to residence time would be L-shaped.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/EvennessAndTurnover_vs_ResTime.png',
      width = 1500, height = 1200, res=200, bg = "white")

p1 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$e.var, 
                           group=sim.data$sim, colour=sim.data$sim)) +
  scale_color_gradient(high="cyan") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="evenness") +
  labs(title="Species evenness") +
  guides(colour=FALSE) + 
  scale_x_log10() +
  coord_fixed(ratio = 2.5)

p2 <- ggplot(sim.data, aes(x=h.tau, y=sim.data$species.turnover, 
                           group=sim.data$sim, colour=sim.data$sim)) +
  scale_color_gradient(high="chartreuse") +
  geom_smooth(mapping=NULL, data=NULL, se=F, stat="smooth") +
  labs(x= expression(tau), y="turnover") +
  labs(title="Species turnover") +
  guides(colour=FALSE) +
  scale_x_log10() +
  coord_fixed(ratio = 2.5)

multiplot(p1, p2, cols=2)
dev.off()
```

## We hypothesized that aspects of diversity would be related to total abundance (N), reflecting the feasible set and the influence of residence time.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")

lm_eqn <- function(df){
    m <- lm(log(y) ~ log(x), df)
    a = format(coef(m)[1], digits = 2) 
    b = format(coef(m)[2], digits = 2) 
    r2 = format(summary(m)$r.squared, digits = 3)
    bquote('S = ' ~ .(a) ~ N^{.(b)})
    }

png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/Diversity_vs_N.png',
      width = 1500, height = 1200, res=200, bg = "white")

df <- data.frame(x=sim.data$total.abundance, y=sim.data$species.richness)
p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$species.richness)) +
  labs(title=lm_eqn(df)) +
  geom_point(colour="snow3") +
  geom_smooth(method='lm',colour="lawngreen", size=2) +
  labs(x= "Total abundance", y="Species richness") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.6)

df <- data.frame(x=sim.data$total.abundance, y=sim.data$e.var)
p2 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$e.var)) +
  labs(title=lm_eqn(df)) +
  geom_point(colour="snow3") +
  geom_smooth(method='lm', colour="lawngreen", size=2) +
  labs(x= "Total abundance", y="Species evenness") +
  scale_x_log10() +
  scale_y_log10() +
  ylim(0.01,1) +
  coord_fixed(ratio = 4.3)

multiplot(p1, p2, cols=2)
dev.off()
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")

lm_eqn <- function(df){
    m <- lm(log(y) ~ log(x), df, na.action=na.exclude)
    a = format(coef(m)[1], digits = 2) 
    b = format(coef(m)[2], digits = 2) 
    r2 = format(summary(m)$r.squared, digits = 3)
    bquote('S = ' ~ .(a) ~ N^{.(b)} ~.(r2))
    }

png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/Diversity_vs_N2.png',
      width = 1500, height = 1200, res=200, bg = "white")

df <- data.frame(x=sim.data$total.abundance, y=sim.data$N.max)
p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$N.max)) +
  labs(title=lm_eqn(df)) +
  geom_point(colour="snow3") +
  geom_smooth(method='lm',colour="lawngreen", size=2) +
  labs(x= "Total abundance", y="Nmax") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1)

multiplot(p1, cols=1)
dev.off()
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=7}
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")

png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/Diversity_vs_N_ByMotion.png',
      width = 1500, height = 1200, res=200, bg = "white")

p1 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$species.richness,
             group=sim.data$sim, colour=sim.data$motion)) +
  geom_point() +
  labs(x= "Total abundance", y="Species richness") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.55)

p2 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$e.var,
                           group=sim.data$sim, colour=sim.data$motion)) +
  geom_point() +
  labs(x= "Total abundance", y="Species evenness") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 3.8)

p3 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$N.max,
                           group=sim.data$sim, colour=sim.data$motion)) +
  geom_point() +
  labs(x= "Total abundance", y="Dominance") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1)

p4 <- ggplot(sim.data, aes(x=sim.data$total.abundance, y=sim.data$skew,
                           group=sim.data$sim, colour=sim.data$motion)) +
  geom_point() +
  labs(x= "Total abundance", y="Rarity") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.4)

multiplot(p1, p2, p3, p4, cols=2)
dev.off()
```