### 2015-09-11 analyses/ForPaper/PrimaryPredictions.Rmd
added more simulation data; generated new figures -KL

---
title: "Residence time, a master variable of ecology, physiology, and natural selection"
author: "Ken Locey, Jay Lennon"
date: "June, 2015"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/hydrobide") 
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
#install.packages("moments")
require("moments")
library(ggplot2)
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Loading files with plotting and diversity functions
source("~/GitHub/hydrobide/tools/Rbin/multiplot.R")
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Load Data
 dat <- read.csv("~/GitHub/hydrobide/results/simulated_data/2015_September/7_Sept/SimData.csv")  
dat$h.tau <- (dat$height * dat$width)/dat$flow.rate
dat$total.abundance[dat$total.abundance<=0] <- 1
dat$species.richness[dat$species.richness<=0] <- 1

dat$e.var[dat$e.var =="NaN"] <- 1
dat$simpson.e[dat$simpson.e =="NaN"] <- 1

f.dat <- subset(dat, motion == "fluid")
r.dat <- subset(dat, motion == "random_walk")
```

## We hypothesized that total abundance (N) and species richness (S) shared humped-shaped relationshiped with residence time, reflecting chemostat theory.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("ISLR")
require(ISLR)

fit <- lm(log10(f.dat$total.abundance) ~ poly(log10(f.dat$h.tau), 2, raw=TRUE))
f.N.2.r2 <- summary(fit)$adj.r.squared
f.N.2.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(log10(f.dat$total.abundance) ~ poly(log10(f.dat$h.tau), 3, raw=TRUE))
f.N.3.r2 <- summary(fit)$adj.r.squared
f.N.3.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(log10(r.dat$total.abundance) ~ poly(log10(r.dat$h.tau), 2, raw=TRUE))
r.N.2.r2 <- summary(fit)$adj.r.squared
r.N.2.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(log10(r.dat$total.abundance) ~ poly(log10(r.dat$h.tau), 3, raw=TRUE))
r.N.3.r2 <- summary(fit)$adj.r.squared
r.N.3.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(log10(f.dat$species.richness) ~ poly(log10(f.dat$h.tau), 2, raw=TRUE))
f.S.2.r2 <- summary(fit)$adj.r.squared
f.S.2.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(log10(f.dat$species.richness) ~ poly(log10(f.dat$h.tau), 3, raw=TRUE))
f.S.3.r2 <- summary(fit)$adj.r.squared
f.S.3.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(log10(r.dat$species.richness) ~ poly(log10(r.dat$h.tau), 2, raw=TRUE))
r.S.2.r2 <- summary(fit)$adj.r.squared
r.S.2.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(log10(r.dat$species.richness) ~ poly(log10(r.dat$h.tau), 3, raw=TRUE))
r.S.3.r2 <- summary(fit)$adj.r.squared
r.S.3.pVal <- anova(fit)$'Pr(>F)'[1]
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/NandS_vs_ResTime.png',
      width = 1500, height = 1200, res=200, bg = "white")

p1 <- ggplot()+
  geom_smooth(aes(x=f.dat$h.tau, y=f.dat$total.abundance, group=f.dat$sim), colour='skyblue1',
    data=f.dat, se=F) +  
  geom_smooth(aes(x=r.dat$h.tau, y=r.dat$total.abundance, group=r.dat$sim), colour='seagreen1',
    data=r.dat, se=F) +

  stat_smooth(aes(x=dat$h.tau, y=dat$total.abundance, group=dat$motion, colour=dat$motion),
    data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
    formula = y ~ poly(x,2), se=FALSE) +

  scale_colour_manual(values = c("navy", "darkgreen"),
      labels = c(paste("Fluid, r^2 = ",round(f.N.2.r2,2)),
      paste("Non-fluid, r^2 = ",round(r.N.2.r2,2))),
      name = "") +

  labs(x= expression(tau), y="Individuals") +
  labs(title="Total abundance") +
  scale_x_log10() +
  scale_y_log10() +
  theme(legend.position="top") +
  coord_fixed(ratio = .50)

p2 <- ggplot()+
  geom_smooth(aes(x=f.dat$h.tau, y=f.dat$species.richness, group=f.dat$sim), colour='skyblue1',
    data=f.dat, se=F) +  
  geom_smooth(aes(x=r.dat$h.tau, y=r.dat$species.richness, group=r.dat$sim), colour='seagreen1',
    data=r.dat, se=F) +

  stat_smooth(aes(x=dat$h.tau, y=dat$species.richness, group=dat$motion, colour=dat$motion),
    data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
    formula = y ~ poly(x,2), se=FALSE) +

  scale_colour_manual(values = c("navy", "darkgreen"),
      labels = c(paste("Fluid, r^2 = ",round(f.S.2.r2,2)),
      paste("Non-fluid, r^2 = ",round(r.S.2.r2,2))),
      name = "") +

  labs(x= expression(tau), y="Number of species") +
  labs(title="Species richness") +
  theme(legend.position="top") +
  scale_x_log10() +
  scale_y_log10() +
  coord_fixed(ratio = 1.0)

multiplot(p1, p2, cols=2)
dev.off()
```

## We hypothesized that the realtionship species evenness (E) to residence time would be U-shaped, and that the relationship of species turnover (B) to residence time would be L-shaped.

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
fit <- lm(f.dat$e.var ~ poly(log10(f.dat$h.tau), 2, raw=TRUE))
f.ev.2.r2 <- summary(fit)$adj.r.squared
f.ev.2.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(f.dat$e.var ~ poly(log10(f.dat$h.tau), 3, raw=TRUE))
f.ev.3.r2 <- summary(fit)$adj.r.squared
f.ev.3.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(r.dat$e.var ~ poly(log10(r.dat$h.tau), 2, raw=TRUE))
r.ev.2.r2 <- summary(fit)$adj.r.squared
r.ev.2.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(r.dat$e.var ~ poly(log10(r.dat$h.tau), 3, raw=TRUE))
r.ev.3.r2 <- summary(fit)$adj.r.squared
r.ev.3.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(f.dat$species.turnover ~ poly(log10(f.dat$h.tau), 2, raw=TRUE))
f.to.2.r2 <- summary(fit)$adj.r.squared
f.to.2.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(f.dat$species.turnover ~ poly(log10(f.dat$h.tau), 3, raw=TRUE))
f.to.3.r2 <- summary(fit)$adj.r.squared
f.to.3.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(r.dat$species.turnover ~ poly(log10(r.dat$h.tau), 2, raw=TRUE))
r.to.2.r2 <- summary(fit)$adj.r.squared
r.to.2.pVal <- anova(fit)$'Pr(>F)'[1]

fit <- lm(r.dat$species.turnover ~ poly(log10(r.dat$h.tau), 3, raw=TRUE))
r.to.3.r2 <- summary(fit)$adj.r.squared
r.to.3.pVal <- anova(fit)$'Pr(>F)'[1]
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE, fig.width=7, fig.height=4}
png(filename = '~/GitHub/hydrobide/results/figs/ForPaper/EvennessAndTurnover_vs_ResTime.png', width = 1500, height = 1200, res=200, bg = "white")

p1 <- ggplot()+
  geom_smooth(aes(x=f.dat$h.tau, y=f.dat$e.var, group=f.dat$sim), colour='skyblue1',
    data=f.dat, se=F) +  
  geom_smooth(aes(x=r.dat$h.tau, y=r.dat$e.var, group=r.dat$sim), colour='seagreen1',
    data=r.dat, se=F) +

  stat_smooth(aes(x=dat$h.tau, y=dat$e.var, group=dat$motion, colour=dat$motion),
    data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
    formula = y ~ poly(x,2), se=FALSE) +

  scale_colour_manual(values = c("navy", "darkgreen"),
      labels = c(paste("Fluid, r^2 = ",round(f.ev.2.r2,2)),
      paste("Non-fluid, r^2 = ",round(r.ev.2.r2,2))),
      name = "") +

  labs(x= expression(tau), y="Evar") +
  labs(title="Species evenness") +
  scale_x_log10() +
  ylim(0,1) +
  theme(legend.position="top") +
  coord_fixed(ratio = 2.0)

p2 <- ggplot()+
  geom_smooth(aes(x=f.dat$h.tau, y=f.dat$species.turnover, group=f.dat$sim), colour='skyblue1',
    data=f.dat, se=F) +  
  geom_smooth(aes(x=r.dat$h.tau, y=r.dat$species.turnover, group=r.dat$sim), colour='seagreen1',
    data=r.dat, se=F) +

  stat_smooth(aes(x=dat$h.tau, y=dat$species.turnover, group=dat$motion, colour=dat$motion),
    data=dat, method = "lm", geom="smooth", linetype="dashed", size=1,
    formula = y ~ poly(x,2), se=FALSE) +

  scale_colour_manual(values = c("navy", "darkgreen"),
      labels = c(paste("Fluid, r^2 = ",round(f.to.2.r2,2)),
      paste("Non-fluid, r^2 = ",round(r.to.2.r2,2))),
      name = "") +

  labs(x= expression(tau), y="Whittaker's") +
  labs(title="Species turnover") +
  scale_x_log10() +
  ylim(0,1) +
  theme(legend.position="top") +
  coord_fixed(ratio = 2.0)

multiplot(p1, p2, cols=2)
dev.off()
```