### 2015-10-05 results/TestResults/SpeedMemory.Rmd
modified model.py, edited Methods.md and Results.md, Added time and memory analysis file and data, added example .Rmd code (Exploratory.Rmd), deleted unecessary .Rmd files, added new simulation data, modified metrics.R and multiplot.R files -KL

---
title: "Speed and Time tests"
author: "Ken Locey"
date: "October, 2015"
header-includes: - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
rm(list=ls()) 
getwd() 
setwd("~/GitHub/simplex") 
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
#install.packages("vegan") 
require("vegan") 
#install.packages("moments")
require("moments")
library(ggplot2)
```

```{r, results = 'hide', echo=FALSE, message = FALSE, warning = FALSE}
## Load Data
dat <- read.csv("~/GitHub/simplex/results/TestResults/SpeedMemory.csv")

sem <- function(x) sqrt(var(x)/length(x))

#is.nan(val) == FALSE
Ns <- dat$N
mn = mean(Ns)
se <- sem(Ns)
print(c(mn, se))

times <- dat$time
mn = mean(times)
se <- sem(times)
print(c(mn, se))

mems <- dat$memory
mn = mean(mems)
se <- sem(mems)
print(c(mn, se))
```